A Logical Solution of the Linearity Challenge

### 2025-12-02 paper/Context.lagda.tex
restored description of Process

\begin{code}[hide]
infix 4 _≃_+_ _≃_,_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  •   : [] ≃ [] + []
  <_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  >_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

@@

It is often the case that the context $\ContextD$ in a splitting
$\Splitting\ContextC\ContextD\ContextE$ is a singleton list $[A]$. For this
particular case, we introduce a dedicated notation that allows us to represent
this case in a more compact and readable way, as
$\SimpleSplitting\ContextC{A}\ContextE$.

\begin{code}
_≃_,_ : Context → Type → Context → Set
Γ ≃ A , Δ = Γ ≃ [ A ] + Δ
\end{code}
\begin{code}[hide]
+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l • = refl
+-empty-l (> p) = cong (_ ∷_) (+-empty-l p)

+-sing-l : ∀{A B Γ} → [ A ] ≃ B , Γ → A ≡ B × Γ ≡ []
+-sing-l (< •) = refl , refl
\end{code}

@@

\begin{code}[hide]
module _ where
\end{code}
\begin{code}
  splitting-example₂ : (A ∷ B ∷ C ∷ D ∷ []) ≃ (B ∷ []) + (A ∷ C ∷ D ∷ [])
  splitting-example₂ = > < ≫
\end{code}

>>>>

\begin{code}[hide]
infix 4 _≃_+_ _∋_⊳_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  •   : [] ≃ [] + []
  <_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  >_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

@@

It is often the case that the context $\ContextD$ in a splitting
$\Splitting\ContextC\ContextD\ContextE$ is a singleton list $[A]$. We use this
particular splitting to describe the presence of a channel of type $A$ in
$\ContextC$ and giving a name $\ContextE$ to the residual context. For this
particular case, we introduce a dedicated notation that allows us to write this
case in a more compact and meaningful way, as
$\SimpleSplitting\ContextC{A}\ContextE$.

\begin{code}
_∋_⊳_ : Context → Type → Context → Set
Γ ∋ A ⊳ Δ = Γ ≃ [ A ] + Δ
\end{code}
\begin{code}[hide]
+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l • = refl
+-empty-l (> p) = cong (_ ∷_) (+-empty-l p)
\end{code}

@@

\begin{code}[hide]
module _ where
\end{code}
\begin{code}
  splitting-example₂ : (A ∷ B ∷ C ∷ D ∷ []) ∋ B ⊳ (A ∷ C ∷ D ∷ [])
  splitting-example₂ = > < ≫
\end{code}

### 2025-12-02 paper/Process.lagda.tex
restored description of Process

CO-DE BRUIJN INDICES

We adopt a \emph{nameless, intrinsically typed} representation of processes,
that is: a channel is identified simply by the presence of its type in the
typing context; a process is also a proof derivation showing that the process is
well typed. This choice increases the effort in the definition of the datatypes
for processes and their operational semantics, but pays off in the rest of the
formalization for a number of reasons:
\begin{itemize}
\item we use context splitting not only to separate resources used by parallel
  processes (\cf \ForkRule and \CutRule), but also to witness that a channel is
  present in the typing context;
\item we conflate processes and typing rules in the same datatype, thus reducing
  the overall number of datatypes defined in the formalization;
\item we embed the typing preservation results within the very definition of the
  operational semantics of processes instead of proving them separately
  (\Cref{sec:agda-semantics}).
\end{itemize}

Our nameless modeling of channels is known in the literature as the \emph{co-de
Bruijn} representation~\cite{McBride18}. The proof that the type of a channel is
found within a typing context essentially encodes the de Bruijn index of the
channel in the typing context of a given process. However, the typing context
only contains types for those channels that actually occur in the process.
Context splitting makes sure that channels are appropriately (and above all
linearly) distributed among sub-processes, so that each channel occurs (and is
used) exactly once.\footnote{Strictly speaking, the typing context of a
\AgdaInductiveConstructor{fail} process may contain channels that do not occur
(and therefore are not used) by the process, but we consider this case as an
exception that does not alter in a substantial way the provided intuition
underlying the use of co-de Bruijn representation.}

We now illustrate the definition of the \AgdaDatatype{Process} datatype and
describe its key aspects in detail. The datatype is indexed by a typing context:

\begin{code}
data Process : Context → Set where
  link      : ∀{A Γ} → Γ ≃ A , [ dual A ] → Process Γ
  fail      : ∀{Γ Δ} → Γ ≃ ⊤ , Δ → Process Γ
  wait      : ∀{Γ Δ} → Γ ≃ ⊥ , Δ → Process Δ → Process Γ
  close     : Process [ 𝟙 ]
  case      : ∀{A B Γ Δ} → Γ ≃ A & B , Δ →
              Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
  select    : ∀{A B Γ Δ} (x : Bool) → Γ ≃ A ⊕ B , Δ →
              Process ((if x then A else B) ∷ Δ) → Process Γ
  join      : ∀{A B Γ Δ} → Γ ≃ A ⅋ B , Δ → Process (B ∷ A ∷ Δ) → Process Γ
  fork      : ∀{A B Γ Δ Γ₁ Γ₂} → Γ ≃ A ⊗ B , Δ → Δ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
  all       : ∀{A Γ Δ} → Γ ≃ `∀ A , Δ →
              ((B : Type) → Process (subst [ B /_] A ∷ Δ)) → Process Γ
  ex        : ∀{A B Γ Δ} → Γ ≃ `∃ A , Δ → Process (subst [ B /_] A ∷ Δ) → Process Γ
  server    : ∀{A Γ Δ} → Γ ≃ `! A , Δ → Un Δ → Process (A ∷ Δ) → Process Γ
  client    : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process (A ∷ Δ) → Process Γ
  weaken    : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process Δ → Process Γ
  contract  : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process (`? A ∷ `? A ∷ Δ) → Process Γ
  cut       : ∀{A Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (dual A ∷ Γ₂) → Process Γ
\end{code}

A \AgdaInductiveConstructor{link} process is well typed in a typing context with
exactly two types $A$ and $B$ which must be related by duality. We use a proof
$p$ that $\Context \simeq [A] + [B]$ instead of requiring $\Context$ to be
simply $[A,B]$ so that it becomes straightforward to define \SLink using
\AgdaFunction{dual-symm} and \AgdaFunction{+-comm}.

% \begin{code}
%    fail      : ∀{Γ Δ} (p : Γ ≃ ⊤ , Δ) → Process Γ
% \end{code}

A \AgdaInductiveConstructor{fail} process simply requires the presence of $\Top$
in the typing context.

% \begin{code}
%    close     : Process  [ 𝟙 ]
% \end{code}

A \AgdaInductiveConstructor{close} process is well typed in the singleton
context $[\One]$.

% \begin{code}
%    wait      : ∀{Γ Δ} (p : Γ ≃ ⊥ , Δ) → Process Δ → Process Γ
% \end{code}

A \AgdaInductiveConstructor{wait} process is well typed in any context
$\ContextC$ that contains $\Bot$ and whose residual $\ContextD$ allows the
continuation to be well typed.

% \begin{code}
%    select    : ∀{A B Γ Δ} (x : Bool) (p : Γ ≃ A ⊕ B , Δ) →
%                Process ((if x then A else B) ∷ Δ) → Process Γ
% \end{code}

We represent the tags $\InTag_1$ and $\InTag_2$ with Agda's boolean values
\AgdaInductiveConstructor{true} and \AgdaInductiveConstructor{false}. In this
way we can use a single constructor \AgdaInductiveConstructor{select} for the
selections $\Select\x{\InTag_i}.P$. According to the semantics of \Calculus, the
channel on which the selection is performed is consumed and the continuation
process uses a fresh continuation channel. As a consequence, the type of the
continuation channel, which is either $A$ or $B$ depending on the value of the
tag $x$, is \emph{prepended} in front of the typing context used to type the
continuation process.

% \begin{code}
%    case      : ∀{A B Γ Δ} (p : Γ ≃ A & B , Δ) →
%                Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
% \end{code}

A \AgdaInductiveConstructor{case} process has two possible continuations. Note
again that the continuation channel, which is received along with the tag, has
either type $A$ or type $B$ depending on the tag and its type is prepended to
the typing context of the continuation process since it is the newest channel.

% \begin{code}
%    fork      : ∀{A B Γ Δ Γ₁ Γ₂} (p : Γ ≃ A ⊗ B , Δ) (q : Δ ≃ Γ₁ + Γ₂) →
%                Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
% \end{code}

A \AgdaInductiveConstructor{fork} process creates two fresh continuation
channels and spaws two continuation processes. Since context splitting is a
ternary relation, we need two splitting proofs $p$ and $q$ to isolate the type
$A \Ten B$ of channel on which the pair of continuation channels is sent and to
distribute the remaining channel across the two continuation processes.

% \begin{code}
%    join      : ∀{A B Γ Δ} (p : Γ ≃ A ⅋ B , Δ) → Process (B ∷ A ∷ Δ) →
%                Process Γ
% \end{code}

The \AgdaInductiveConstructor{join} process has a single continuation which uses
the pair of continuation channels received from a channel of type $A \Par B$.
Note that $B$ is prepended in front of $A$, somewhat implying that the second
continuation is bound \emph{after} the first even though, from a technical
standpoint, both continuations are bound simultaneously. We could have prepended
the types $A$ and $B$ also in the opposite order, provided that the reduction
rule \RFork (formalized in \Cref{sec:reduction-agda}) is suitably adjusted.

% \begin{code}
%    server    : ∀{A Γ Δ} (p : Γ ≃ `! A , Δ) (un : Un Δ) →
%                Process (A ∷ Δ) → Process Γ
% \end{code}

A \AgdaInductiveConstructor{server} constructor introduces a channel of type
$\OfCourse A$ given a continuation process that consumes a channel of type $A$
and a proof $\mathit{un}$ that every other channel in $\ContextD$ is
unrestricted.\Luca{Introdurre terminologia e dire che unrestricted vuol dire che
il canale può essere ``duplicato'' e scartato.}

% \begin{code}
%    client    : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process (A ∷ Δ) → Process Γ
%    weaken    : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process Δ → Process Γ
%    contract  : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process (`? A ∷ `? A ∷ Δ) →
%                Process Γ
% \end{code}

The constructors \AgdaInductiveConstructor{client},
\AgdaInductiveConstructor{weaken} and \AgdaInductiveConstructor{contract}
introduce a channel of type $\WhyNot A$ and follow the same pattern of the
previous process forms. Note that, in the case of
\AgdaInductiveConstructor{contract}, the contraction of $\WhyNot A$ applies to
the first two channels in the typing context of the continuation. This is
consistent with the fact that \Calculus uses an explicit prefix
$\Contract\x\y\z$ that models contraction and binds $y$ and $z$ in the
continuation process. In a pen-and-paper presentation, like the one by
Wadler~\citep{Wadler14}, it is arguably more convenient to have weaking and
contraction typing rules without the corresponding process forms.

% \begin{code}
%    cut       : ∀{A Γ Γ₁ Γ₂} (p : Γ ≃ Γ₁ + Γ₂) →
%                Process (A ∷ Γ₁) → Process (dual A ∷ Γ₂) → Process Γ

%    all       : ∀{A Γ Δ} (p : Γ ≃ `∀ A , Δ) →
%                ((B : Type) → Process (subst [ B /_] A ∷ Δ)) → Process Γ
%    ex        : ∀{A B Γ Δ} (p : Γ ≃ `∃ A , Δ) →
%                Process (subst [ B /_] A ∷ Δ) → Process Γ
% \end{code}

A \AgdaInductiveConstructor{cut} process incorporates a duality proof for the
peer endpoints of the channel being restricted and a splitting proof that
distributes the free channels among the two parallel processes.

@@

\begin{code}
↭process : ∀{Γ Δ} → Γ ↭ Δ → Process Γ → Process Δ
\end{code}
\begin{code}[hide]
↭process π (link p) with ↭solo π p
... | Δ′ , q , π′ rewrite ↭solo-inv π′ = link q
↭process π (fail p) with ↭solo π p
... | Δ′ , q , π′ = fail q
↭process π (wait p P) with ↭solo π p
... | Δ′ , q , π′ = wait q (↭process π′ P)
↭process π close rewrite ↭solo-inv π = close
↭process π (case p P Q) with ↭solo π p
... | Δ′ , q , π′ = case q (↭process (prep π′) P) (↭process (prep π′) Q)
↭process π (select x p P) with ↭solo π p
... | Δ′ , q , π′ = select x q (↭process (prep π′) P)
↭process π (join p P) with ↭solo π p
... | Δ′ , q , π′ = join q (↭process (prep (prep π′)) P)
↭process π (fork p q P Q) with ↭solo π p
... | Δ′ , p′ , π′ with ↭split π′ q
... | Δ₁ , Δ₂ , q′ , π₁ , π₂ = fork p′ q′ (↭process (prep π₁) P) (↭process (prep π₂) Q)
↭process π (all p P) with ↭solo π p
... | Δ' , q , π' = all q λ B → ↭process (prep π') (P B)
↭process π (ex p P) with ↭solo π p
... | Δ' , q , π' = ex q (↭process (prep π') P)
↭process π (server p un P) with ↭solo π p
... | Δ′ , q , π′ = server q (↭un π′ un) (↭process (prep π′) P)
↭process π (client p P) with ↭solo π p
... | Δ′ , q , π′ = client q (↭process (prep π′) P)
↭process π (weaken p P) with ↭solo π p
... | Δ′ , q , π′ = weaken q (↭process π′ P)
↭process π (contract p P) with ↭solo π p
... | Δ′ , q , π′ = contract q (↭process (prep (prep π′)) P)
↭process π (cut p P Q) with ↭split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = cut q (↭process (prep π₁) P) (↭process (prep π₂) Q)
\end{code}

>>>>

We adopt a \emph{nameless, intrinsically-typed} representation of processes,
whereby a channel is identified by the position of its type in the typing
context and a process is also a proof derivation showing that the process is
well typed. This choice increases the effort in the definition of the datatypes
for defining processes and their operational semantics, but pays off in the rest
of the formalization for a number of reasons:
\begin{itemize}
\item we use context splitting not only to separate resources used by parallel
  processes (\cf \ForkRule and \CutRule), but also to witness that a channel is
  present in the typing context;
\item we conflate processes and typing rules in the same datatype, thus reducing
  the overall number of datatypes defined in the formalization;
\item we embed the typing preservation results within the very definition of the
  operational semantics of processes instead of proving them separately
  (\Cref{sec:agda-semantics}).
\end{itemize}

Our nameless representation of channels is known in the literature as the
\emph{co-de Bruijn} representation of identifiers~\cite{McBride18}. The proof
that the type of a channel is found within a typing context essentially encodes
the de Bruijn index of the channel in the typing context of a given process.
However, the typing context only contains types for those channels that actually
occur in the process. Context splitting makes sure that channels are
appropriately (and above all linearly) distributed among sub-processes, so that
each channel occurs (and is used) exactly once.\footnote{Strictly speaking, the
typing context of a \AgdaInductiveConstructor{fail} process may contain channels
that do not occur (and therefore are not used) by the process, but we consider
this case as an exception that does not alter in a substantial way the provided
intuition underlying the use of co-de Bruijn representation.}

We now illustrate the definition of the \AgdaDatatype{Process} datatype and then
describe its key aspects in detail. The datatype is indexed by a typing context
$\Context$ according to which the process is well typed:

\begin{code}
data Process : Context → Set where
  link      : ∀{A Γ} → Γ ≃ [ A ] + [ dual A ] → Process Γ
  fail      : ∀{Γ Δ} → Γ ∋ ⊤ ⊳ Δ → Process Γ
  wait      : ∀{Γ Δ} → Γ ∋ ⊥ ⊳ Δ → Process Δ → Process Γ
  close     : Process [ 𝟙 ]
  case      : ∀{A B Γ Δ} → Γ ∋ A & B ⊳ Δ →
              Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
  left      : ∀{A B Γ Δ} → Γ ∋ A ⊕ B ⊳ Δ → Process (A ∷ Δ) → Process Γ
  right     : ∀{A B Γ Δ} → Γ ∋ A ⊕ B ⊳ Δ → Process (B ∷ Δ) → Process Γ
  join      : ∀{A B Γ Δ} → Γ ∋ A ⅋ B ⊳ Δ → Process (B ∷ A ∷ Δ) → Process Γ
  fork      : ∀{A B Γ Δ Γ₁ Γ₂} → Γ ∋ A ⊗ B ⊳ Δ → Δ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
  all       : ∀{A Γ Δ} → Γ ∋ `∀ A ⊳ Δ →
              ((X : Type) → Process (subst [ X /_] A ∷ Δ)) → Process Γ
  ex        : ∀{A B Γ Δ} → Γ ∋ `∃ A ⊳ Δ → Process (subst [ B /_] A ∷ Δ) → Process Γ
  server    : ∀{A Γ Δ} → Γ ∋ `! A ⊳ Δ → Un Δ → Process (A ∷ Δ) → Process Γ
  client    : ∀{A Γ Δ} → Γ ∋ `? A ⊳ Δ → Process (A ∷ Δ) → Process Γ
  weaken    : ∀{A Γ Δ} → Γ ∋ `? A ⊳ Δ → Process Δ → Process Γ
  contract  : ∀{A Γ Δ} → Γ ∋ `? A ⊳ Δ → Process (`? A ∷ `? A ∷ Δ) → Process Γ
  cut       : ∀{A Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (dual A ∷ Γ₂) → Process Γ
\end{code}

Let us start from the constructors representing input/output actions, namely
those from \AgdaInductiveConstructor{fail} through
\AgdaInductiveConstructor{contract}. They all basically follow the same pattern:
their first (explicit) argument is a splitting of the form
$\SimpleSplitting\ContextC{C}\ContextD$ that identifies the channel on which the
process is acting within the typing context $\ContextC$ and where the shape of
$C$ is determined by the typing rule; the subsequent arguments are terms
corresponding to the continuations of the process being represented, each
indexed by its own context. Crucially, the type of the continuation channels are
\emph{prepended} to the context that indexes the continuation processes using
the \AgdaInductiveConstructor{\_∷\_} list constructor. This is consistent with
the fact that the continuation channels have been freshly introduced, and the
channel on which the process operates has been consumed.
%
There are just a few deviations from this pattern, which we discuss next.

The constructor \AgdaInductiveConstructor{close} lacks a splitting argument:
since a process $\Close\x$ is well typed in a singleton context $x : \One$, it
is easier to specify such context directly as the index of
\AgdaDatatype{Process} rather than through a splitting.

The constructor \AgdaInductiveConstructor{fork}, which represents a process
$\Fork\x\y\z{P}{Q}$, has an additional splitting argument
$\Splitting\ContextD{\Context_1}{\Context_2}$ which distributes the resources in
$\ContextD$ among the two continuation processes $P$ and $Q$.

The constructor \AgdaInductiveConstructor{all}, which represents a process
$\Forall\x\X\y.P$, models the continuation $P$ using higher-order abstract
syntax (HOAS~\cite{}): the continuation $P$ is produced by a function applied to
a \AgdaInductiveConstructor{Type} $X$. Notice how the constructor enforces the
side condition of \ForallRule \emph{by definition}:
%
$A$ has type
$\AgdaDatatype{PreType}~(\AgdaInductiveConstructor{suc}~\AgdaInductiveConstructor{zero})$,
that is a pre-type with \emph{at most one} free type variable (that is $X$),
whereas $\ContextD$ is a list of $\AgdaDatatype{Type} =
\AgdaDatatype{PreType}~\AgdaInductiveConstructor{zero}$, that is a list of
pre-types without free type variables. Therefore, $X$ cannot occur in
$\ContextD$ as required by \ForallRule.

Finally, the constructor \AgdaInductiveConstructor{server}, which represents a
process $\Server\x\y.P$, has an additional argument of type
$\AgdaDatatype{Un}~\ContextD$ enforcing the property that $\ContextD$ is
unrestricted.

The two remaining constructors, \AgdaInductiveConstructor{link} and
\AgdaInductiveConstructor{cut}, deal with links and cuts in the expected way. A
link is well typed in a context
$\Splitting\Context{[A]}{[\AgdaFunction{dual}~A]}$. We use a splitting instead
of requiring $\Context$ to be $[A,\AgdaFunction{dual}~A]$ so that it becomes
straightforward to define \SLink using \AgdaFunction{+-comm}.
%
The constructor \AgdaInductiveConstructor{cut} models a process $\Cut\x{P}{Q}$
and uses splitting to distribute the resources in $\Context$ among $P$ and $Q$.

@@

\begin{code}
↭process : ∀{Γ Δ} → Γ ↭ Δ → Process Γ → Process Δ
\end{code}
\begin{code}[hide]
↭process π (link p) with ↭solo π p
... | Δ′ , q , π′ rewrite ↭solo-inv π′ = link q
↭process π (fail p) with ↭solo π p
... | Δ′ , q , π′ = fail q
↭process π (wait p P) with ↭solo π p
... | Δ′ , q , π′ = wait q (↭process π′ P)
↭process π close rewrite ↭solo-inv π = close
↭process π (case p P Q) with ↭solo π p
... | Δ′ , q , π′ = case q (↭process (prep π′) P) (↭process (prep π′) Q)
↭process π (left p P) with ↭solo π p
... | Δ′ , q , π′ = left q (↭process (prep π′) P)
↭process π (right p P) with ↭solo π p
... | Δ′ , q , π′ = right q (↭process (prep π′) P)
↭process π (join p P) with ↭solo π p
... | Δ′ , q , π′ = join q (↭process (prep (prep π′)) P)
↭process π (fork p q P Q) with ↭solo π p
... | Δ′ , p′ , π′ with ↭split π′ q
... | Δ₁ , Δ₂ , q′ , π₁ , π₂ = fork p′ q′ (↭process (prep π₁) P) (↭process (prep π₂) Q)
↭process π (all p P) with ↭solo π p
... | Δ' , q , π' = all q λ B → ↭process (prep π') (P B)
↭process π (ex p P) with ↭solo π p
... | Δ' , q , π' = ex q (↭process (prep π') P)
↭process π (server p un P) with ↭solo π p
... | Δ′ , q , π′ = server q (↭un π′ un) (↭process (prep π′) P)
↭process π (client p P) with ↭solo π p
... | Δ′ , q , π′ = client q (↭process (prep π′) P)
↭process π (weaken p P) with ↭solo π p
... | Δ′ , q , π′ = weaken q (↭process π′ P)
↭process π (contract p P) with ↭solo π p
... | Δ′ , q , π′ = contract q (↭process (prep (prep π′)) P)
↭process π (cut p P Q) with ↭split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = cut q (↭process (prep π₁) P) (↭process (prep π₂) Q)
\end{code}

### 2025-12-02 paper/calculus.tex
restored description of Process

\begin{table}
    \caption{Syntax of \Calculus.}\label{tab:syntax}
    \begin{tabular}{c}
        \begin{math}
            \displaystyle
            \begin{array}[t]{rcll}
                P, Q & ::= & \Link\x\y            & \text{link} \\
                    & | & \Fail\x                 & \text{empty choice} \\
                    & | & \Wait\x.P               & \text{wait} \\
                    & | & \Close\x                & \text{close} \\
                    & | & \Case\x\z{P}{Q}         & \text{case} \\
                    & | & \Select\x{\InTag_i}\z.P & \text{selection} \\
                    & | & \Join\x\y\z.P           & \text{join} \\
                    & | & \Fork\x\y\z{P}{Q}       & \text{fork} \\
                    & | & \Forall\x\X\y.P         & \text{type input} \\
                    & | & \Exists\x{A}\y.P        & \text{type output} \\
                    & | & \Server\x\y.P           & \text{server accept} \\
                    & | & \Client\x\y.P           & \text{client request} \\
                    & | & \Weaken\x.P             & \text{weakening} \\
                    & | & \Contract\x\y\z.P       & \text{contraction} \\
                    & | & \Cut[A]\x{P}{Q}         & \text{cut} \\
            \end{array}
        \end{math}
    \end{tabular}
\end{table}

>>>>

\begin{table}
    \caption{Syntax of \Calculus.}\label{tab:syntax}
    \begin{tabular}{c}
        \begin{math}
            \displaystyle
            \begin{array}[t]{rcll}
                P, Q & ::= & \Link\x\y            & \text{link} \\
                    & | & \Fail\x                 & \text{fail} \\
                    & | & \Wait\x.P               & \text{wait} \\
                    & | & \Close\x                & \text{close} \\
                    & | & \Case\x\z{P}{Q}         & \text{case} \\
                    & | & \Select\x{\InTag_i}\z.P & \text{left/right selection} \\
                    & | & \Join\x\y\z.P           & \text{join} \\
                    & | & \Fork\x\y\z{P}{Q}       & \text{fork} \\
                    & | & \Forall\x\X\y.P         & \text{for all} \\
                    & | & \Exists\x{A}\y.P        & \text{exists} \\
                    & | & \Server\x\y.P           & \text{server} \\
                    & | & \Client\x\y.P           & \text{client} \\
                    & | & \Weaken\x.P             & \text{weakening} \\
                    & | & \Contract\x\y\z.P       & \text{contraction} \\
                    & | & \Cut[A]\x{P}{Q}         & \text{cut} \\
            \end{array}
        \end{math}
    \end{tabular}
\end{table}
